
```{r}
library("gapminder")
library("dplyr")

gapminder

```


Filtering for one year

```{r}

gapminder %>%
  filter(year == 1957)

```


Filtering for one year and country

```{r}

gapminder %>%
  filter(year == 1957, country == "China")

```


Arranging observations by life expectancy

```{r}
gapminder %>%
  arrange(lifeExp)

# Sort in descending order of lifeExp

gapminder %>%
  arrange(desc(lifeExp))

```


Filtering and arranging

```{r}

gapminder %>%
  filter(year == 1957) %>%
  arrange(desc(pop))

```


Using mutate to change or create a column

```{r}
# use mutate to change lifeExp to be in mouths

gapminder %>%
  mutate(lifeExp = 12 * lifeExp)

# use mutate to create a new column called lifeExpMonths
gapminder %>%
  mutate(lifeExpMonths = 12 * lifeExp)

```


Combining filter, mutate and arrange

```{r}
gapminder %>%
  filter(year == 2007) %>%
  mutate(lifeExpMonths = 12 * lifeExp) %>%
  arrange(desc(lifeExpMonths))

```


Visualizing with ggplot2

```{r}

library(ggplot2)

gapminder_1952 <- gapminder %>%
  filter(year == 1952)

ggplot(gapminder_1952, aes(x = pop, y = gdpPercap)) +
  geom_point()

```

```{r}

gapminder_1952 <- gapminder %>%
  filter(year == 1952)

ggplot(gapminder_1952, aes(x = pop, y = lifeExp)) +
  geom_point()

```


Putting the x-axis on a log scale

```{r}

gapminder_1952 <- gapminder %>%
  filter(year == 1952)

ggplot(gapminder_1952, aes(x = pop, y = lifeExp)) +
  geom_point() +
  scale_x_log10()

```


Putting the x-axis and y-axis on a log scale

```{r}

ggplot(gapminder_1952, aes(x = pop, y = gdpPercap)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10()

```


Adding color to a scatter plot

```{r}

ggplot(gapminder_1952, aes(x = pop, y = lifeExp, color = continent)) +
  geom_point() +
  scale_x_log10()

```


```{r}

ggplot(gapminder_1952, aes(x = pop, y = lifeExp, color = continent, size = gdpPercap)) +
  geom_point() +
  scale_x_log10()

```


Faceting

Creating a subgraph for each continent

```{r}

gapminder_1952

ggplot(gapminder_1952, aes(x = pop, y = lifeExp)) +
  geom_point() +
  scale_x_log10() +
  facet_wrap(~ continent)

```


Faceting by year

```{r}

ggplot(gapminder, aes(x = gdpPercap, y = lifeExp, color = continent, size = pop)) +
  geom_point() +
  scale_x_log10() +
  facet_wrap(~ year)

```
